---
title: "Yield_planting_exploration_sea"
Author: Michelle Yiv
format: html
embed-resources: true
---

## Setup

```{r, message=FALSE, warning=FALSE}
# Load libraries
library(tidyverse)
library(here)
library(janitor)
library(dbplyr)
library(duckdb)
library(DBI)
library(wesanderson)
```

```{r, message=FALSE}
# Read in data
calendar <- read_csv(here('data', 'crop_yields.csv')) 
yield <- read_csv(here('data','FAOSTAT_data_en_4-1-2025 (1).csv'))
```

## Data exploration and cleaning

```{r}
# View column names for crop yields
#colnames(yield)

# Note that there are periods instead of underscores
# Possibly join on Area/Location column? Different naming conventions for some area Viet nam vs. Vietnam(South)
```

```{r}
# View column names for crop calendar
#colnames(calendar)
```

```{r}
# Data cleaning for calendar

calendar_cleaned <- calendar %>%
  
  # Rename to match with yield dataset
  rename(Area = Location) %>%
  
  clean_names()

# Replace . in column names to _
names(calendar_cleaned) <- ifelse(str_detect(names(calendar_cleaned), "\\."),
                    str_replace_all(names(calendar_cleaned), "\\.", "_"),
                    names(calendar_cleaned))
```

```{r}
# Cleaning for crop yield

yield_cleaned <- yield %>%
  
  # Rename area code column for clarity
  rename(area_code = `Area Code (M49)`) %>%
  
  # Rename item code column for clarity
  rename(item_code = `Item Code (CPC)`) %>%
  
  # Change to lowercase
  clean_names() 
```

### Focus on planting and harvesting ranges, items, and yield by South East Asian countries

```{r}
# Referring to schema diagram, select columns of interest

calendar_table <- calendar_cleaned %>%
  select(c("data_id", "area", "crop","plant_start_date", "plant_end_date", "plant_range", "harvest_start_date", "harvest_end_date", "harvest_range"))

yield_table <- yield_cleaned %>%
  select(c("item", "item_code", "value", "unit", "year", "element", "area")) %>%
  filter(element == "Yield")

country_table <- yield_cleaned %>%
  select(c("area_code", "area"))
```

```{r}
yield_table_sea <- yield_table %>%
  filter(area %in% c("Malaysia", "Indonesia", "Laos", "Myanmar", "Singapore", "Cambodia", "Philippines", "Thailand", "Brunei", "East Timor", "Vietnam", "Nepal")) %>%
  filter(year == 2022) %>% 
  rename(crop = item)
  
calendar_table_sea <- calendar_table %>%
  filter(area %in% c("Malaysia", "Indonesia", "Laos", "Myanmar", "Singapore", "Cambodia", "Philippines", "Thailand", "Brunei", "East Timor", "Vietnam", "Nepal"))
```

```{r}
# Save to CSV
# write.csv(yield_table_sea, file = "yield.csv", row.names = FALSE)
# write.csv(calendar_table_sea, file = "calendar.csv", row.names = FALSE)
```

```{r, message=FALSE}
# Find average for planting and harvest ranges
calendar_table_sea_avg <- calendar_table_sea %>%
  group_by(area, crop) %>%
  summarize(
    avg_plant_range = mean(plant_range, na.rm = TRUE),
    avg_harvest_range = mean(harvest_range, na.rm = TRUE)
  ) %>%
  ungroup()
```

```{r}
# Join tables
avg_cal_yield <- left_join(calendar_table_sea_avg, yield_table_sea, by = c("crop", "area")) %>%
  drop_na()

```

```{r}
# Visualization 
ggplot(avg_cal_yield, aes(x = reorder(crop, -value), y = value, fill = area)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "2022 Crop Yield in Selected South East Asian Countries",
    x = "",
    y = "Yield (kg/ha)",
    fill = "Country"
  ) +
  theme_minimal() +
  scale_fill_manual(values = rev(wes_palette("Cavalcanti1"))) +
  theme(text = element_text(size = 24))

# ggsave("crop.png", plot1, bg = NULL, height = 14, width = 12)
```

```{r}
## Looking at crops other than potatoes, reference only
avg_cal_yield %>%
  filter(crop != "Potatoes") %>%
ggplot(aes(x = reorder(crop, -value), y = value, fill = area)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "2022 Crop Yield in Selected South East Asian Countries",
    x = "",
    y = "Yield (kg/ha)",
    fill = "Country"
  ) +
  theme_minimal() +
  scale_fill_manual(values = rev(wes_palette("Cavalcanti1"))) +
  theme(text = element_text(size = 24))
```

## Results part 1
Here, we can see that Myanmar and the Philippines produce a substantial amount of potatoes compared to the other countries. This is likely due to these countries having cool, highland regions that are optimal for potato cultivation. Philippines leads again when looking at rice production, although Nepal and Cambodia have similar yields. Interestingly, Thailand has the least amount of rice yield despite it being heavily marketed. OVerall, Nepal has the most yield compared to the other countries, leading the wheat and millet crops while still being a top producer for rice.


## Connect to database

```{r}
# Connect to database
conn <- DBI::dbConnect(duckdb::duckdb(), dbdir="database.db")

# Ensure tables are present
DBI::dbListTables(conn)
```

```{r}
# Look at tables
calendar_db <- tbl(conn, "calendar")
yield_db <- tbl(conn, "yield")

#calendar_db
#yield_db
```

```{r}
# Join tables
yield_cal_df <- left_join(calendar_db, yield_db, by=c("area", "crop")) %>%
  collect()

# Drop NA values
yield_cal_df <- yield_cal_df %>%
  drop_na() ## These are from having planting information but no yield information. Since we focus on rice, we can drop these values.
```

```{r}
# Visualize planting and harvest times for rice
 yield_cal_df %>%
  filter(crop == "Rice") %>%
  mutate(
    
    # Add years to dates
    plant_start_date = mdy(paste0(plant_start_date, "/2022")),
    plant_end_date = mdy(paste0(plant_end_date, "/2022")),
    

    harvest_start_date = mdy(paste0(harvest_start_date, "/2022")),
    harvest_end_date = mdy(paste0(harvest_end_date, "/2022")),
    
    
    # Move harvest date to after plant date. Assumption that harvest is AFTER planting to fix plot confusion
    harvest_start_date = if_else(harvest_start_date < plant_start_date,
                                harvest_start_date + years(1),
                                harvest_start_date),
    
    harvest_end_date = if_else(harvest_end_date < plant_start_date,
                              harvest_end_date + years(1),
                              harvest_end_date)
  ) %>%
  ggplot() +
  
  # Planting period
  geom_segment(aes(x = plant_start_date, xend = plant_end_date, y = area, yend = area),
               color = "#33a02c", size = 4) +
  # Harvest period
  geom_segment(aes(x = harvest_start_date, xend = harvest_end_date, y = area, yend = area),
               color = "#fdbf6f", size = 4) +
  labs(
    title = "Rice Planting & Harvest Periods by Country (2022-2023)",
    x = "Date",
    y = "Country"
  ) +
  theme_minimal() + scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  theme(text = element_text(size = 24))

# ggsave("timeline.png", plot2, width = 12, height = 8)
```
Here, we look at the average planting and harvesting times for the four rice producing countries. There are intersting patterns here, which may be due to how data was handled. Thailand seems to have the most relatively even planting and harvest period. The Philippines has a relatively short planting period and an extended harvest period, while Cambodia has the opposite. 
```{r, message=FALSE, warning=FALSE}
# Close connection
DBI::dbDisconnect(conn, shutdown= TRUE)
```
